Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2019-05-31 doc/network.tex
stuff

\begin{figure}[!pbth]
\includegraphics[width=\textwidth]{../fig/individual_and_population.pdf}
\caption{\textbf{Comparison of individual- and population-level kernels.}
(Left) an individual-level kernel of an infected individual with latent period of 11.4 days followed by infectious period of 5 days. 
(Right) a population-level kernel of infected individuals with latent and infectious periods exponentially distributed with means of 11.4 and 5 days, respectively. 
Shaded areas under the curves are equal to individual- and population-level reproductive numbers, both of which are set to 2 in this example.
}
\label{fig:indpop}

@@

\begin{figure}[!pbth]
\includegraphics[width=\textwidth]{../fig/temporal_effect.pdf}
\caption{\textbf{Temporal variation in the mean observed generation interval.}
A deterministic Susceptible-Exposed-Infectious-Recovered (SEIR) model was simulated using Ebola-like parameters. The backward and the censored mean generation interval were calculated over the course of an epidemic.
The dotted horizontal line represents the mean intrinsic generation interval.
}
\label{fig:censor}
\end{figure}

@@

\jd{I want to keep thinking about network choice. Would a random network be useful? Are we careful enough about time windows (exponential phase vs (roughly) full single epidemic?}
\swp{I might have an idea... let me think about this a bit more...}
\begin{figure}[!pbth]
\includegraphics[width=\textwidth]{../fig/local_effect.pdf}
\caption{
\textbf{Spatial effects on realized generation intervals.}
Theoretical distributions and means are shown in color (and are the same in each panel, for reference). Simulated distributions and means are shown in black.
(Left) the intrinsic generation-interval distribution corresponds to all contacts by a focal individual, regardless of whether the individual contacted is susceptible (simulated on a star network).
(Middle) the egocentric generation-interval distribution corresponds to the distribution of all infectious contacts by the focal individual with susceptible individuals, in the case where the focal individual is the only possible infector (simulated on a star network).
(Right) realized generation-interval distributions are shorter than egocentric distributions in general, because contacts can be wasted when susceptibles become infected through other routes (simulated on a homogeneous network).
All figures were generated using 5000 stochastic simulations on a network with 5 nodes (1 infector and 4 susceptibles).
}
\label{fig:local}
\end{figure}

@@

When generation intervals are sampled through contact tracing, there will be four effects present in the sample: (1) right-censoring effect, (2) egocentric depletion effect, (3) local depletion effect and (4) global depletion effect.
We can correct explicitly for the egocentric effect and, in the case of exponential growth, the right-censoring effect.
While the other two effects are difficult to measure, we can make qualitative predictions about their effects on the realized generation intervals and reproductive numbers: 
both local and global depletion effects reduce number of infections that occur and shorten generation intervals.
% SWP: I would rather put these in Discussion?
% but the time scale on which they matter differs:
% local depletion will be important early in an epidemic whereas global depletion will be small until sufficient amount of susceptible individuals become infected.
% As a result, we expect the initial spread of a disease and the realized reproductive number to be heavily dependent on local structure of the contact network near the source of an outbreak.

@@

\begin{figure}[!pbth]
\includegraphics[width=\textwidth]{../fig/cmp_reproductive.pdf}
\caption{\textbf{Comparison of estimates of reproductive number based on various methods.}
Using the observed generation-interval distributions without correcting for right-censoring severely underestimates the reproductive number.
Similarly, using intrinsic and egocentric generation-interval distributions without accounting for local spatial effects overestimates the reproductive number.
Both population-level and individual-level methods provide estimates of reproductive number that are consistent with the empirical estimates, which we define as the average number of secondary cases generated by the first 100 infected individuals.
Boxplots are generated using 100 stochastic simulations of the SEIR model on an empirical network using Ebola-like parameters.
}
\label{fig:cmp}
\end{figure}

>>>>

\begin{figure}[!pbth]
\includegraphics[width=\textwidth]{../fig/individual_and_population.pdf}
\caption{\textbf{Comparison of individual- and population-level kernels.}
(Left) an individual-level kernel of an infected individual with latent period of 11.4 days followed by infectious period of 5 days. 
(Right) a population-level kernel of infected individuals with latent and infectious periods exponentially distributed with means of 11.4 and 5 days, respectively. 
Shaded areas under the curves are equal to individual- and population-level reproductive numbers, both of which are set to 2 in this example.
Parameters are chosen to reflect the West African Ebola outbreak \citep{who2014ebola}.
}
\label{fig:indpop}

@@

\begin{figure}[!pbth]
\includegraphics[width=\textwidth]{../fig/temporal_effect.pdf}
\caption{\textbf{Temporal variation in the mean observed generation interval.}
A deterministic Susceptible-Exposed-Infectious-Recovered (SEIR) model was simulated using Ebola-like parameters \citep{who2014ebola}: mean latent period $1/\sigma = 5 \textrm{ days}$, mean infectious period $1/\gamma = 11.4 \textrm{ days}$, and the basic reproductive number $\RR_0 = 2$. 
The backward and the censored mean generation interval were calculated over the course of an epidemic.
The dotted horizontal line represents the mean intrinsic generation interval.
}
\label{fig:censor}
\end{figure}

@@

\jd{I want to keep thinking about network choice. Would a random network be useful? Are we careful enough about time windows (exponential phase vs (roughly) full single epidemic?}
\swp{I might have an idea... let me think about this a bit more...}
\begin{figure}[!pbth]
\includegraphics[width=\textwidth]{../fig/local_effect.pdf}
\caption{
\textbf{Spatial effects on realized generation intervals.}
Theoretical distributions and means are shown in color (and are the same in each panel, for reference). Simulated distributions and means are shown in black.
(Left) the intrinsic generation-interval distribution corresponds to all contacts by a focal individual, regardless of whether the individual contacted is susceptible (simulated on a star network).
(Middle) the egocentric generation-interval distribution corresponds to the distribution of all infectious contacts by the focal individual with susceptible individuals, in the case where the focal individual is the only possible infector (simulated on a star network).
(Right) realized generation-interval distributions are shorter than egocentric distributions in general, because contacts can be wasted when susceptibles become infected through other routes (simulated on a homogeneous network).
All figures were generated using 5000 stochastic simulations on a network with 5 nodes (1 infector and 4 susceptibles) with Ebola-like parameters \citep{who2014ebola}:
mean latent period $1/\sigma = 5 \textrm{ days}$ and mean infectious period $1/\gamma = 11.4 \textrm{ days}$. 
Per-pair contact rate $\lambda = 0.25 \textrm{ days}^{-1}$ is chosen to be sufficiently high so that the differences between generation-interval distributions are clear.
}
\label{fig:local}
\end{figure}

@@

When generation intervals are sampled through contact tracing, there will be four effects present in the sample: (1) right-censoring effect, (2) egocentric depletion effect, (3) local depletion effect and (4) global depletion effect.
We can correct explicitly for the egocentric effect and, in the case of exponential growth, the right-censoring effect.
While the other two effects are difficult to measure, we can make qualitative predictions about their effects on the realized generation intervals and reproductive numbers: 
both local and global depletion effects reduce number of infections that occur and shorten generation intervals.

@@

\begin{figure}[!pbth]
\includegraphics[width=\textwidth]{../fig/cmp_reproductive.pdf}
\caption{\textbf{Comparison of estimates of reproductive number based on various methods.}
Using the observed generation-interval distributions without correcting for right-censoring severely underestimates the reproductive number.
Similarly, using intrinsic and egocentric generation-interval distributions without accounting for local spatial effects overestimates the reproductive number.
Both population-level and individual-level methods provide estimates of reproductive number that are consistent with the empirical estimates, which we define as the average number of secondary cases generated by the first 100 infected individuals.
Boxplots are generated using 100 stochastic simulations of the SEIR model on an empirical network using Ebola-like parameters \citep{who2014ebola}: mean latent period $1/\sigma = 5 \textrm{ days}$ and mean infectious period $1/\gamma = 11.4 \textrm{ days}$. 
Per-pair contact rate $\lambda = 0.08 \textrm{ days}^{-1}$ is chosen to be sufficiently high such that differences are clear. \swp{Appendix}
}
\label{fig:cmp}
\end{figure}
